---
title: "MacLearn_Data_PeerAssessment"
author: "Adesh Shrestha(www.emacshop.com)"
date: "January 13, 2016"
output: html_document
---

<b><h2>Synopsis</b></h2>

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <a href="http://groupware.les.inf.puc-rio.br/har">http://groupware.les.inf.puc-rio.br/har</a> (see the section on the Weight Lifting Exercise Dataset).

<b><h2>Synopsis</b></h2>

Training data can be downloaded from <a href="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv">Training Data</a>

Testing data can be downloaded from <a href="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv">Testing Data</a>

<b><h2>Expected Output</b></h2>

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.

<ol type="1">
  <li>Your submission should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).</li>
  <li>You should also apply your machine learning algorithm to the 20 test cases available in the test data above.Please submit your predictions in appropriate format to the programming assignment for automated grading. See the programming assignment for additional details.</li>
</ol>  

<b><h2>Detailed Code File</b></h2>

Kindly review the Analysis.R file in the gitHub repo to review the code developed to complete the assignment.

<b><h2>Load Required Packages</b></h2>

I have assumed that the packages are already installed on the machine. In case if packages are not already installed, please use require("package name") or install.packages("package name") to load the package in your machine. Thanks..

<b><h3>Load caret, rpart, rpart.plot, RColorBrewer, rattle, randomForest packages</b></h3>

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r}
library(caret)
```

```{r}
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(rpart)
```

```{r}
library(randomForest)
```

<b><h3>Set seed to reproduce same results</b></h3>

```{r}
set.seed(12345)
```

<b><h3>Loading data in machine cache</b></h3>

For this assignment, data has been loaded in cache. However, in case if user wants to load the data in the hard disc before proceeding, generic options are available to do the same.

```{r}
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(url(trainUrl), na.strings=c("NA","#DIV/0!",""))
testing <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
```

<b><h3>Partitioning the dataset into training and testing</b></h3>

Based on the lecture notes, below code is used to create training and testing dataset. 60% data is considered for the training set and remaining 40% is considered for testing the model.

A line of code is written to view the dimensions of the dataset.

```{r}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ]; myTesting <- training[-inTrain, ]
dim(myTraining)
dim(myTesting)
```